Update and evaluation of the ozone dry deposition in the OsloCTM3 v1.0

### 2018-12-07 DryDep.tex
Spell-checked, corrected version with respect to velocity computation.

wherein aerodynamic $R_a$, quasi-laminar layer $R^i_b$, and canopy resistances $R^i_c$ are key components \citep{AE:Wesely1989,ACP:Seinfeld2006}. For all gases, $R_a$ is the same, while $R^i_b$ and $R^i_c$ vary from gas to gas and also depend on land surface types (e.g., ice/snow, water, urban, desert, agricultural land, deciduous forest, coniferous forest etc.). Originally, \citet{AE:Wesely1989} used fixed seasonal average dry deposition resistances for each land surface type. For all three types of resistances in this Wesely-type parameterization, more process-oriented formulations have been developed and validated over the years. \citet{ACP:Luhar2017} have validated ozone dry deposition to the ocean with respect to three different formulations of surface resistances. An update on the ozone surface resistance over snow and ice covered surfaces has been provided from combined model and observation studies \citep[][$v^\chem{O_3}_\text{ice/snow} = 1/10000\,\unit{m\,s^{-1}}$]{ACP:Helmig2007}. Canopy conductance is parameterized at the single-leaf-level (stomatal conductance) for various plant function types (PFT) as well as for single plant species based on empirical studies \citep{PTRS:Jarvis1976, BallBerry1987, ACP:Simpson2012, ICP:MappingManual2017}. But progress has also been made on process oriented modeling of stomatal conductance \citep{AFM:Anderson2000,PP:Buckley2017}. The variety of differing formulations and choices of parameters leads to a wide spread of results in model intercomparisons \citep{ACP:Hardacre2015,AE:Derwent2018} and about $20\,\unit{\%}$ uncertainty on the resulting total dry deposition \citep{ACP:Monks2015}.\\

@@

@@

Finally, we take a look at the different global as well as hemispheric sinks for ozone (Table~\ref{tab:ozone_sinks}). Despite its vastness, the ocean amounts only to about $23\,\unit{\%}$ of the global ozone sink in the OsloCTM3 with operative EMEP scheme, while ice and snow account for far less than $1\,\unit{\%}$. The remainder is deposited to various other land surfaces of which deserts might be the most neglected in process modeling. The total annual dry deposition in the EMEP scheme is about $\frac{1}{3}$ below the multi-model-mean result by \citet{ACP:Hardacre2015}. But also the results of ~\citet{ACP:Luhar2017} yield a $(19-27)\,\unit{\%}$ lower ozone dry deposition than the models participating in the model intercomparison, with deposition to ocean ranging between $(12-21)\,\unit{\%}$.
%
\begin{table*}[t]
\caption{Total ozone dry deposition for the respective year in \unit{Tg\,a^{-1}} separated into ocean, ice and, land contributions.}
\begin{tabular}{lccccccccccccr}
\tophline
\multirow{3}{*}{Simulation} & \multicolumn{3}{c}{Ocean} & \multicolumn{3}{c}{Ice} & \multicolumn{3}{c}{Land} & \multicolumn{3}{c}{Total} & $\Delta^\dagger$\\
& NH & SH & Global & NH & SH & Global & NH & SH & Global & NH & SH & Global\\
& \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & (\unit{\%})\\
\middlehline
Wesely\_type & 88.7 & 110.3 & 199.0 & 0.7 & 5.0 & 5.7 & 346.1 & 153.5 & 499.6 & 612.4 & 347.9 & 960.2 & 46.8\\
EMEP\_full & 67.1 & 84.7 & 151.8 & 0.6 & 4.4 & 5.0 & 236.0 & 112.3 & 348.2 & 408.6 & 245.6 & 654.2 & 0.0\\
EMEP\_offLight & 66.6 & 84.1 & 150.7 & 0.6 & 4.4 & 5.0 & 267.8 & 129.8 & 397.6 & 451.3 & 267.2 & 718.6 & 9.8\\
EMEP\_offPhen & 67.1 & 84.6 & 151.7 & 0.6 & 4.4 & 5.0 & 239.9 & 116.0 & 355.9 & 413.9 & 249.9 & 663.7 & 1.5\\
EMEP\_SWVL4 & 68.1 & 86.1 & 154.2 & 0.6 & 4.5 & 5.1 & 241.8 & 115.0 & 356.8 & 417.2 & 250.7 & 667.9 & 2.1\\
EMEP\_ppgs & 67.1 & 84.6 & 151.7 & 0.6 & 4.4 & 5.0 & 238.7 & 112.6 & 351.3 & 411.4 & 246.0 & 657.4 & 0.5\\
EMEP\_ppgssh & 67.0 & 84.5 & 151.6 & 0.6 & 4.4 & 5.0 & 243.9 & 119.4 & 363.3 & 419.4 & 253.6 & 673.0 & 2.9\\
EMEP\_ppgssh\_ice & 66.5 & 84.8 & 151.3 & 0.2 & 1.3 & 1.4 & 240.3 & 120.4 & 360.7 & 413.0 & 251.1 & 664.1 & 1.5\\
EMEP\_ppgs\_2005 & 66.8 & 83.5 & 150.4 & 0.6 & 4.4 & 5.0 & 237.8 & 110.0 & 347.8 & 409.2 & 240.7 & 649.9 & -0.7\\
\bottomhline
\end{tabular}
\belowtable{$^\dagger$: Difference of global annual total relative to \emph{EMEP\_full}.}% Table Footnotes
\label{tab:ozone_sinks}
\end{table*}
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comparison with observations}
\label{subsec:obs}
In this section, we compare our results to observations. For all comparisons, we use the original resolution of the OsloCTM3 instead of the re-gridded.

In Fig.~\ref{fig:mmm_drydep_stations}a), seasonal cycles of average ozone dry deposition fluxes for six observation sites are shown. We have computed a model average for all sensitivity studies at the closest grid point and show the $1\,\sigma$ error band. The shaded area around the multi-model-mean indicates the broad range of model results but is not an actual error band since such is not given in \citet{ACP:Hardacre2015}. At about $4$ of $6$ sites, the EMEP scheme performs better than the Wesely scheme and better than or the same as the multi-model-mean. We use a $\chi^2$-test
\begin{equation}
  \chi^2 = \sum_{i=1}^{12}\frac{\left(\overline{\chem{O_3}}_{\text{DD,\,}i}^\text{sim}-\overline{\chem{O_3}}_{\text{DD,\,}i}^\text{obs}\right)^2}{\sigma_i^2},
\end{equation}
with an estimated standard deviation of observation $\sigma_i=1\,\unit{mmol\,m^{-2}\,s^{-1}}$ and divide it by the number degrees of freedom (NDF) to assess this subjective analysis in a more objective way. The closer to $1$ this test scores, the better does the simulation represent the observation. A score between $0$ and $1$ indicates that the estimated $\sigma$ is too small. The results of the $\chi^2$-test are shown together with the divergences in Fig.~\ref{fig:mmm_drydep_stations}. The $\chi^2$-test reveals also that in $4$ of $6$ cases the EMEP scheme improves the performance of the OsloCTM3 with respect to ozone dry deposition fluxes, although a satisfying result is only achieved for two sites (Castel Porziano, Blodgett Forest).
With only one full year of simulation, the model uncertainty regarding the seasonal cycle at observational sites cannot be quantitatively assessed. Further more, the observational averages uses at most nine years worth of data which questions the statistical reliability of these data. Among other aspects that might affect the model performance the horizontal as well vertical resolution play an important role. In the rather coarse resolution, good matches between observation and model are only to be expected if the station is at a location which is representative for an area similar to the model resolution.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conclusions[Discussion and conclusions]  %% \conclusions[modified heading if necessary]
\label{sec:conc}
We have presented an update of the dry deposition scheme in the OsloCTM3 from prescribed dry deposition velocities \citep{AE:Wesely1989,JGR:Hough1991} to a more process oriented parameterization taking the state of the atmosphere and vegetation into account \citep{ACP:Simpson2012}. The dry deposition to vegetation is now sensitive to parameters such as photosynthetic active radiation, water vapor pressure deficit, soil water content, and plant phenology. The update decreases the annual amount of ozone dry deposition up to $100\,\unit{\%}$ especially over large desert regions and the oceans. Compared to results from a multi-model evaluation \citep{ACP:Hardacre2015}, the total annual ozone dry deposition in the OsloCTM3 is about $33\,\unit{\%}$ below the average. However, there seems to be a tendency, that the newer TF~HTAP models show a lower total annual dry deposition of ozone than older models, indicating that newer developments in the models lead to decreasing ozone dry deposition and increasing tropospheric ozone burden \citep[e.g.,][]{ACP:Luhar2017,AE:Hu2017}.

Most of the decrease in ozone dry deposition in the OsloCTM3 is due to changes in dry deposition velocities over the ocean and deserts. In case of desert and grasslands this change is at the order of one magnitude. Over the ocean it is due to the accumulative effect of a small absolute change in dry deposition which is especially amplified in the southern hemisphere. With respect to available measurements of dry deposition velocities of ozone over desert \citep{AE:Gusten1995} and ocean \citep{JGR:Helmig2012}, the new OsloCTM3 dry deposition scheme slightly underestimates ozone dry deposition velocities over the former and overestimate them over the latter. Regarding the vastness of the ocean and the ongoing desertification, it may be worthwhile to study dry deposition in these regimes at a more process-oriented level, e.g., 2-layer gas exchange with ocean waters \citep{ACP:Luhar2017}, wave braking and spray \citep{ACP:Pozzer2006}, or influence of desert dust.

>>>>

wherein aerodynamic $R_a$, quasi-laminar layer $R^i_b$, and canopy resistances $R^i_c$ are key components \citep{AE:Wesely1989,ACP:Seinfeld2006}. For all gases, $R_a$ is the same, while $R^i_b$ and $R^i_c$ vary from gas to gas and also depend on land surface types (e.g., ice/snow, water, urban, desert, agricultural land, deciduous forest, coniferous forest etc.). Originally, \citet{AE:Wesely1989} used fixed seasonal average dry deposition resistances for each land surface type. For all three types of resistances in this Wesely-type parameterization, more process-oriented formulations have been developed and validated over the years. \citet{ACP:Luhar2017} have validated ozone dry deposition to the ocean with respect to three different formulations of surface resistances. An update on the ozone surface resistance over snow and ice covered surfaces has been provided from combined model and observation studies \citep[][$v^\chem{O_3}_\text{ice/snow} = 1/10000\,\unit{m\,s^{-1}}$]{ACP:Helmig2007}. Canopy conductance is parameterized at the single-leaf-level (stomatal conductance) for various plant function types (PFT) as well as for single plant species based on empirical studies \citep{PTRS:Jarvis1976, BallBerry1987, ACP:Simpson2012, ICP:MappingManual2017}. But progress has also been made on process-oriented modeling of stomatal conductance \citep{AFM:Anderson2000,PP:Buckley2017}. The variety of differing formulations and choices of parameters leads to a wide spread of results in model intercomparisons \citep{ACP:Hardacre2015,AE:Derwent2018} and about $20\,\unit{\%}$ uncertainty on the resulting total dry deposition \citep{ACP:Monks2015}.\\

@@

There seems to be a discrepancy between the OsloCTM3 response and the multi-model-mean, since the Wesely scheme is closer to the multi-model-mean with respect to total annual ozone dry deposition, while the reconstructed $v_\text{DD}^\chem{O_3}$ of the EMEP scheme match better. Without knowing the actual $[\chem{O_3}(z_0)]$ of the TF~HTAP models that participated in the model intercomparison, we cannot elaborate on this discrepancy.

@@

Finally, we take a look at the different global as well as hemispheric sinks for ozone (Table~\ref{tab:ozone_sinks}). Despite its vastness, the ocean amounts only to about $23\,\unit{\%}$ of the global ozone sink in the OsloCTM3 with operative EMEP scheme, while ice and snow account for far less than $1\,\unit{\%}$. The remainder is deposited to various other land surfaces of which deserts might be the most neglected in process-modeling. The total annual dry deposition in the EMEP scheme is about $\frac{1}{3}$ below the multi-model-mean result by \citet{ACP:Hardacre2015}. But also the results of ~\citet{ACP:Luhar2017} yield a $(19-27)\,\unit{\%}$ lower ozone dry deposition than the models participating in the model intercomparison, with deposition to ocean ranging between $(12-21)\,\unit{\%}$.
%
\begin{table*}[t]
\caption{Total ozone dry deposition for the respective year in \unit{Tg\,a^{-1}} separated into ocean, ice and, land contributions.}
\begin{tabular}{lccccccccccccr}
\tophline
\multirow{3}{*}{Simulation} & \multicolumn{3}{c}{Ocean} & \multicolumn{3}{c}{Ice} & \multicolumn{3}{c}{Land} & \multicolumn{3}{c}{Total} & $\Delta^\dagger$\\
& NH & SH & Global & NH & SH & Global & NH & SH & Global & NH & SH & Global\\
& \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & (\unit{\%})\\
\middlehline
Wesely\_type & 88.7 & 110.3 & 199.0 & 0.7 & 5.0 & 5.7 & 346.1 & 153.5 & 499.6 & 612.4 & 347.9 & 960.2 & 46.8\\
EMEP\_full & 67.1 & 84.7 & 151.8 & 0.6 & 4.4 & 5.0 & 236.0 & 112.3 & 348.2 & 408.6 & 245.6 & 654.2 & 0.0\\
EMEP\_offLight & 66.6 & 84.1 & 150.7 & 0.6 & 4.4 & 5.0 & 267.8 & 129.8 & 397.6 & 451.3 & 267.2 & 718.6 & 9.8\\
EMEP\_offPhen & 67.1 & 84.6 & 151.7 & 0.6 & 4.4 & 5.0 & 239.9 & 116.0 & 355.9 & 413.9 & 249.9 & 663.7 & 1.5\\
EMEP\_SWVL4 & 68.1 & 86.1 & 154.2 & 0.6 & 4.5 & 5.1 & 241.8 & 115.0 & 356.8 & 417.2 & 250.7 & 667.9 & 2.1\\
EMEP\_ppgs & 67.1 & 84.6 & 151.7 & 0.6 & 4.4 & 5.0 & 238.7 & 112.6 & 351.3 & 411.4 & 246.0 & 657.4 & 0.5\\
EMEP\_ppgssh & 67.0 & 84.5 & 151.6 & 0.6 & 4.4 & 5.0 & 243.9 & 119.4 & 363.3 & 419.4 & 253.6 & 673.0 & 2.9\\
EMEP\_ppgssh\_ice & 66.5 & 84.8 & 151.3 & 0.2 & 1.3 & 1.4 & 240.3 & 120.4 & 360.7 & 413.0 & 251.1 & 664.1 & 1.5\\
EMEP\_ppgs\_2005 & 66.8 & 83.5 & 150.4 & 0.6 & 4.4 & 5.0 & 237.8 & 110.0 & 347.8 & 409.2 & 240.7 & 649.9 & -0.7\\
\bottomhline
\end{tabular}
\belowtable{$^\dagger$: Difference of global annual total relative to \emph{EMEP\_full}.}% Table Footnotes
\label{tab:ozone_sinks}
\end{table*}
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comparison with observations}
\label{subsec:obs}
In this section, we compare our results to observations. For all comparisons, we use the original resolution of the OsloCTM3 ($2.25^\circ\times 2.25^\circ$) instead of the re-gridded ($3^\circ\times 3^\circ$). Although we do not assess the impacts of differing resolution in models on compatibility with observation, we can assume that both, high and low biases, may occur.

In Fig.~\ref{fig:mmm_drydep_stations}a), seasonal cycles of average ozone dry deposition fluxes for six observation sites are shown. We have computed a model average for all sensitivity studies at the closest grid point and show the $1\,\sigma$ error band. The shaded area around the multi-model-mean indicates the broad range of model results but is not an actual error band since such is not given in \citet{ACP:Hardacre2015}. At about $4$ of $6$ sites, the EMEP scheme performs better than the Wesely scheme and better than or the same as the multi-model-mean. We use a $\chi^2$-test
\begin{equation}
  \chi^2 = \sum_{i=1}^{12}\frac{\left(\overline{\chem{O_3}}_{\text{DD,\,}i}^\text{sim}-\overline{\chem{O_3}}_{\text{DD,\,}i}^\text{obs}\right)^2}{\sigma_i^2},
\end{equation}
with an estimated standard deviation of observation $\sigma_i=1\,\unit{mmol\,m^{-2}\,s^{-1}}$ and divide it by the number degrees of freedom (NDF) to assess this subjective analysis in a more objective way. The closer to $1$ this test scores, the better does the simulation represent the observation. A score between $0$ and $1$ indicates that the estimated $\sigma$ is too small. The results of the $\chi^2$-test are shown together with the divergences in Fig.~\ref{fig:mmm_drydep_stations}b). The $\chi^2$-test reveals also that in $4$ of $6$ cases the EMEP scheme improves the performance of the OsloCTM3 with respect to ozone dry deposition fluxes, although a satisfying result is only achieved for two sites (Castel Porziano, Blodgett Forest).
With only one full year of simulation, the model uncertainty regarding the seasonal cycle at observational sites cannot be quantitatively assessed. Further more, the observational averages comprise at most nine years worth of data. Statistically, these data may still subject to interannual variability. Among other aspects, the horizontal as well as vertical resolution play an important role in the model performance. In the rather coarse resolution, good matches between observation and model are only to be expected if the station is at a location which is representative for an area similar to the model resolution.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conclusions[Discussion and conclusions]  %% \conclusions[modified heading if necessary]
\label{sec:conc}
We have presented an update of the dry deposition scheme in the OsloCTM3 from prescribed dry deposition velocities \citep{AE:Wesely1989,JGR:Hough1991} to a more process-oriented parameterization taking the state of the atmosphere and vegetation into account \citep{ACP:Simpson2012}. The dry deposition to vegetation is now sensitive to parameters such as photosynthetic active radiation, water vapor pressure deficit, soil water content, and plant phenology. Due to changes in dry deposition velocities $v_\text{DD}^\chem{O_3}$, the annual amount of ozone dry deposition decreases by up to $100\,\unit{\%}$ especially over large desert regions and the oceans. Compared to results from a multi-model evaluation \citep{ACP:Hardacre2015}, the total annual ozone dry deposition in the OsloCTM3 is about $33\,\unit{\%}$ below the average. However, there seems to be a tendency, that the newer TF~HTAP models show a lower total annual dry deposition of ozone than older models, indicating that newer developments in the models lead to decreasing ozone dry deposition and increasing tropospheric ozone burden \citep[e.g.,][]{ACP:Luhar2017,AE:Hu2017}. We found the response of the OsloCTM3 to the changes in dry deposition velocities consistent between the old and the new dry deposition scheme. A decrease in $v_\text{DD}^\chem{O_3}$ leads to a decrease in total ozone dry deposition and an increase in surface ozone concentration $[\chem{O_3}(z_0)]$. However, there is an apparent discrepancy when comparing with the multi-model-mean. Without knowing $[\chem{O_3}(z_0)]$ of the TF~HTAP models, this cannot be resolved.

Most of the decrease in ozone dry deposition in the OsloCTM3 is due to changes in dry deposition velocities over the ocean and deserts. In case of desert and grasslands this change is at the order of one magnitude. Over the ocean it is due to the accumulative effect of a small absolute change in dry deposition which is especially amplified in the southern hemisphere. With respect to available measurements of dry deposition velocities of ozone over desert \citep{AE:Gusten1995} and ocean \citep{JGR:Helmig2012}, the new OsloCTM3 dry deposition scheme slightly underestimates ozone dry deposition velocities over the former and overestimate them over the latter. Regarding the vastness of the ocean and the ongoing desertification, it may be worthwhile to study dry deposition in these regimes at a more process-oriented level, e.g., 2-layer gas exchange with ocean waters \citep{ACP:Luhar2017}, wave braking and spray \citep{ACP:Pozzer2006}.

### 2018-12-07 DryDep_supplement.tex
Spell-checked, corrected version with respect to velocity computation.

\documentclass[manuscript]{copernicus}
\usepackage{graphicx}
\usepackage{subcaption}
%\usepackage{subfigure}
\graphicspath{{pictures/}}          % graphics
\begin{document}
\clearpage
\setcounter{page}{1}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Supplement}
\subsection*{S.1 Aerodynamical resistance}
\begin{table}[!htbp]
  \caption{Molecular diffusivity for a gas $i$ given as ration $D_\chem{H_2O}/D_i$, reactivity $f^i_0$, and effective Henry's Law constant $H^i_*$.}
  \begin{tabular}{lcrr@{$\cdot$}l}
    \tophline
    & $D_\chem{H_2O}/D_i$ & $f^i_0$ & \multicolumn{2}{l}{$H^i_*$}\\
    & & & \multicolumn{2}{l}{(\unit{M\,atm^{-1}})}\\
    \middlehline
    \chem{O_3}     & 1.6 & 1.0 & 1 & $10^{-2}$\\
    \chem{SO_2}    & 1.9 & 0.0 & 1 & $10^5$\\
    \chem{NO_2}    & 1.6 & 0.1 & 1 & $10^{-2}$\\
    \chem{H_2O_2}  & 1.4 & 1.0 & 1 & $10^5$\\
    \chem{HCHO}    & 1.3 & 0.0 & 6 & $10^3$\\
    \chem{NO}      & 1.3 & 0.0 & 2 & $10^{-3}$\\
    \chem{CH_3CHO} & 1.6 & 15.0 & \multicolumn{2}{l}{$0$}\\
    \bottomhline
  \end{tabular}
  %\belowtable{} % Table Footnotes
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{S.2 Dynamic viscosity of air and quasi-laminar resistance}
\begin{figure}[!htbp]
  \centering
  \includegraphics[height=0.4\textheight]{pictures/dynamic_viscosity.pdf}
  \caption{Sutherland's law and divergence linear fits with one or two degrees of freedom. The fit $\mu(T) = m\cdot T$ is implemented in the OsloCTM3. Despite its strong divergence at lower and higher temperatures, the impact of this choice on the results is neglicible.}
\end{figure}
%
\begin{figure}[!htbp]
  \centering
  \includegraphics[height=0.4\textheight]{pictures/dynamic_viscosity_zo_Rb.pdf}
  \caption{Comparison of resulting $z_o$ and $R_b$ for $\mu(T) = m\cdot T$  (left panel) and Sutherland's law (right panel). The example displays the distributions for \chem{H_2O} after one day of model integration for both, rough and calm sea case.}
\end{figure}

@@

  \caption{Resulting $R_b$ for different species. Shown are results of one day model integrations for both, Southerland's law and $\mu(T) = m\cdot T$. Shown are also the percentages above / below the given threashold.}

@@

\clearpage
\subsection*{S.4 Latitude dependent vegetation height}
\begin{figure}[!htbp]
  \centering
  \includegraphics[height=0.4\textheight]{pictures/tree_height.pdf}
  \caption{Extension of the latitude dependent vegetation height from northern hemisphere mid latitudes only to a global describtion. Example heights are shown for coniferous, decideous, needleleaf, and broadleaf forests as well as Mediterranean shrub. As reference the original range of EMEP is added.}
\end{figure}

>>>>

\documentclass[manuscript]{copernicus}
\usepackage{graphicx}
\usepackage{subcaption}
%\usepackage{subfigure}
\graphicspath{{pictures/}}          % graphics
\begin{document}
\clearpage
\setcounter{page}{1}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Supplement}
\subsection*{S.1 Aerodynamical resistance}
\begin{table}[!htbp]
  \caption{Molecular diffusivity for a gas $i$ given as ration $D_\chem{H_2O}/D_i$, reactivity $f^i_0$, and effective Henry's Law constant $H^i_*$.}
  \begin{tabular}{lcrr@{$\cdot$}l}
    \tophline
    & $D_\chem{H_2O}/D_i$ & $f^i_0$ & \multicolumn{2}{l}{$H^i_*$}\\
    & & & \multicolumn{2}{l}{(\unit{M\,atm^{-1}})}\\
    \middlehline
    \chem{O_3}     & 1.6 & 1.0 & 1 & $10^{-2}$\\
    \chem{SO_2}    & 1.9 & 0.0 & 1 & $10^5$\\
    \chem{NO_2}    & 1.6 & 0.1 & 1 & $10^{-2}$\\
    \chem{H_2O_2}  & 1.4 & 1.0 & 1 & $10^5$\\
    \chem{HCHO}    & 1.3 & 0.0 & 6 & $10^3$\\
    \chem{NO}      & 1.3 & 0.0 & 2 & $10^{-3}$\\
    \chem{CH_3CHO} & 1.6 & 15.0 & \multicolumn{2}{l}{$0$}\\
    \bottomhline
  \end{tabular}
  %\belowtable{} % Table Footnotes
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{S.2 Dynamic viscosity of air and quasi-laminar resistance}
\begin{figure}[!htbp]
  \centering
  \includegraphics[height=0.4\textheight]{pictures/dynamic_viscosity.pdf}
  \caption{Sutherland's law and divergence linear fits with one or two degrees of freedom. The fit $\mu(T) = m\cdot T$ is implemented in the OsloCTM3. Despite its strong divergence at lower and higher temperatures, the impact of this choice on the results is negligible.}
\end{figure}
%
\begin{figure}[!htbp]
  \centering
  \includegraphics[height=0.4\textheight]{pictures/dynamic_viscosity_zo_Rb.pdf}
  \caption{Comparison of resulting $z_o$ and $R_b$ for $\mu(T) = m\cdot T$  (left panel) and Sutherland's law (right panel). The example displays the distributions for \chem{H_2O} after one day of model integration for both, rough and calm sea case.}
\end{figure}

@@

  \caption{Resulting $R_b$ for different species. Shown are results of one day model integrations for both, Sutherland's law and $\mu(T) = m\cdot T$. Shown are also the percentages above / below the given threshold.}

@@

\clearpage
\subsection*{S.4 Latitude dependent vegetation height}
\begin{figure}[!htbp]
  \centering
  \includegraphics[height=0.4\textheight]{pictures/tree_height.pdf}
  \caption{Extension of the latitude dependent vegetation height from northern hemisphere mid latitudes only to a global description. Example heights are shown for coniferous, deciduous, needleleaf, and broadleaf forests as well as Mediterranean shrub. As reference the original range of EMEP is added.}
\end{figure}
